\scuba: iterative map-making with the Sub-Millimetre User Reduction Facility

### 2012-12-04 smurf_paper.tex
some more minor edits

There are three general styles of map-making that are relevant to
reducing bolometer data in the literature. The simplest ``direct
methods'' involve some basic processing of the data to remove as much
noise contamination as possible, (e.g., using baseline removal and
other simple filters), and then re-gridding these cleaned data into a
map. Such was the basic recipe for the reduction of chopped data from
SCUBA-2's predecessor SCUBA
\citep{1998ASPC..145..216J,2000ASPC..216..559J}, and MAMBO
\citep[e.g.,][]{omont2001}, another camera from the same generation. A
more recent example is the analysis of SPT data
\citep{schaffer2011}. Generally speaking, such methods are fast,
although depending on the science goals and noise properties of the
data, they may not achieve the best noise performance on the angular
scales of interest. A related method that has been popular for
reducing data in fields of faint point sources with Bolocam
\citep[e.g.,][]{laurent2005}, and its younger sibling the Aztronomical
Thermal Emission Camera \citep[AzTEC, e.g.,][]{scott2008}, is
principal component analysis (PCA) cleaning. A statistical
``black-box'' removes the most correlated components of the bolometer
signals, enabling the detection of point-sources close to the
theoretical white-noise limits of the detectors, with reasonable
computation times when small numbers of bolometers are involved
(hundreds rather than thousands of detectors). However, PCA cleaning
is not a good solution for producing maps of extended structures,
since such sources produce correlated signals amongst many detectors,
and are removed by this procedure \citep[an exception is the iterative
PCA approach of][]{aguirre2011}. Furthermore, in the case of \scuba,
performing PCA on even a single subarray (typically $\sim$900
detectors) can be prohibitively slow.

@@

The third approach adopted here for \scuba\ is a compromise between
the previous two methods. Under the assumption that a significant
portion of the (predominantly low-frequency) non-white noise sources
can be modelled, an iterative solution is obtained for both the
astronomical image and the parameters of the noise model. Since the
remaining (non-modelled) noise sources are assumed to be white, a
single scalar \rms\ may be calculated for all of the data points from
a given bolometer to characterise its noise distribution (since the
noise at any instant in time is uncorrelated with others, and with the
data from other bolometers), greatly simplifying the inversion step
that is so complicated in the maximum-likelihood methods. In the
iterative approach memory usage usually scales linearly with $N$, the
product of the number of bolometers, $N_\mathrm{b}$, with the number
of samples in time, $N_\mathrm{t}$. The computation time is the
product of the number of iterations, $N_\mathrm{iter}$, with the time
per iteration, $t_\mathrm{iter}$. For the algorithm described in this
paper, $N_\mathrm{iter}$ is typically in the range $\sim$5--100
(depending on the \snr\ of large-scale structures, and the size of the
map), and $t_\mathrm{iter}$ scales as $N_\mathrm{t} \log
N_\mathrm{t}$, since Fast Fourier Transforms (FFTs) of the time-series
are typically performed. By comparison, the fastest maximum-likelihood
methods that account for inter-bolometer correlations have an
execution time with an $N_\mathrm{b}^2$ dependence, although it is
possible to limit memory usage to be linear in $N_\mathrm{t}$
\citep[e.g.,][]{patanchon2008}.

@@

A reasonable model for correlated noise in \scuba\ is a single
``common-mode'' signal seen by all of the bolometers. Iterative
estimation and removal of this signal significantly lowers the $1/f$
knee, without compromising structures on angular scales smaller than
the array footprint. Residual independent drifts at lower frequencies
are removed with an iterative high-pass filter. This strategy enables
SMURF to reduce data on time scales commensurate with the observing
times (or significantly faster), on single high-end desktop computers
using the . For reference, all of the data analysis in this paper was
performed on a machine with a 64-bit 8-core central processing unit
operating at 2.7\,GHz, and 48\,Gb of memory. SMURF has been
successfully used as part of a real-time pipeline \citep[based on the
Observatory Reduction and Acquisition Control - Data Reduction system,
or ORAC-DR,][]{gibb2005} offering feedback to observers at the
telescope. The pipeline is also used to generate products for the JCMT
Science Archive \citep{2011ASPC..442..203E} hosted by the Canadian
Astronomy Data Centre (CADC). Details of the pipeline design are given
in \citet{gibb2005} and \citet{2008AN....329..295C}.

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{skynoise}
\caption{Illustration of the dominance of atmospheric variations on
  long-timescales using 35\,min. of data taken in opaque, and
  highly-variable weather conditions (line-of-sight 450\,\micron\
  opacity ranging from 12 to 22) from the fully-commissioned
  instrument, observation 18 on 20110324. Red and blue lines show the
  850 and 450\,\micron\ common-mode signals, corresponding to the left-
  and right- vertical axes, respectively. The black line shows the
  line-of-sight opacity from the WVM as a proxy for atmospheric
  emission (arbitrary scale). Bolometer baseline drifts on timescales
  $\gsim$100\,s are clearly correlated with the atmosphere in this
  case.}
\label{fig:skynoise}
\end{figure}

The data shown in Figs.~\ref{fig:bolos_mix} and \ref{fig:pspec} were
taken during fairly stable weather conditions (although with different
450\,\micron\ line-of-sight opacities $\sim$0.9 and 6.5 in the S2SRO
and post-upgrade data sets, respectively), and are dominated at
low-frequencies ($\lsim$2\,Hz) by scan-synchronous noise, rather than
underlying variations in the atmosphere, or ``sky noise''. However,
only $\sim$100\,s of data were analysed. In Fig.~\ref{fig:skynoise},
an order-of-magnitude longer (fully-commissioned) data set is plotted
from observation 18 on 20110324, taken during more opaque, and highly
variable weather conditions (450\,\micron\ line-of-sight opacities
ranging from 12 to 22, scaled from the WVM data which are shown as a
black line). Unlike the earlier plots, it is clear that the bolometer
baseline drifts on timescales $\gsim$100\,s are tightly correlated
with the atmospheric opacity, which serves as a proxy for atmospheric
emission. The weaker correlation between the WVM and the 450\,micron\
bolometers is probably due to the atmosphere being more transmissive
at 850\micron, therefore the emission detected at 450\,\micron\ comes
from water vapour that is closer to the telescope.

@@

As an alternative, we also investigated an algorithm in which the FFT
of $x_i$ is simply truncated to the target sample rate before
transforming back to the time domain (i.e., applying a hard-edged
low-pass filter). In practice the noise performance was
indistinguishable, and required slightly longer execution time than using
the method above. Therefore this method is not used.

@@

SMURF uses FFTs of the bolometer data extensively for filtering. Data
that have been flagged as bad for any reason (for instance, due to the
presence of spikes, steps, or unusual common-mode signal) when need to
be excluded from this FFT. For this reason, each contiguous block of
bad data samples is filled with artificial data before taking the
FFT. A least squares linear fit is performed to the 50 samples
preceding the block, and a similar fit is performed to the 50 samples
following the block. These are used to estimate the expected values at
the start and end of the block of bad values. The bad values are then
replaced by linear interpolation between the expected start and end
values. Gaussian noise is added with a standard deviation equal to the
mean of the RMS residuals in the two fits. These flagged and filled
portions of the data are then given a weight of zero when estimating
the map.

@@

Finally, data that are taken while the telescope is either moving
extremely slowly, or extremely fast, are flagged and ignored when
estimating the map. When the telescope is moving slowly, astronomical
signals appear at low frequencies in the bolometer data that are
dominated by $1/f$ noise (in the extreme case of no motion, the
bolometers have zero sensitivity). When the telescope is moving
significantly faster than 600\,arcsec\,s$^{-1}$, both the thermal
response of the bolometers, and the anti-aliasing filter in the
readout electronics (Section~\ref{sec:bolos}) can significantly
distort signals.  The default allowable range of speed is
30--1000\,arcsec\,s$^{-1}$. These thresholds primarily reject short
periods when the telescope was stationary prior to commencing the scan
pattern (generally speaking the speed is constant once the scan
begins). The scan speed can also briefly exceed the upper limit, even
for a requested speed of 600\,arcsec, if the pointing system is
struggling for large-area scans at high elevation.

@@

In principle, the effects of the bolometer thermal response,
anti-aliasing filter, and data acquisition lags should be accounted
for; together they produce a net lag in the bolometer signals of about
4.5\,ms (Section~\ref{sec:bolos}). This can be modelled by convolving
the ``pure'' bolometer signals with a single system response
function. Similarly, this effect could be removed using de-convolution
as a pre-processing step. The impact of this lag is primarily a
reduction in point-source sensitivity since scans at different
position angles will detect the source at slightly different positions
(and therefore reduce the peak signal once averaged together). This
attenuation is most significant at 450\,\micron\ given the smaller
beam size, and at the greatest scan speeds. For reference, at a speed
of 150\,arcsec\,s$^{-1}$ (typical for CV daisies of calibrators and
deep fields) the attenuation is a negligible $<$1\% at 850\,\micron,
and $\sim$1\% at 450\,\micron. However, at the maximum speed of
600\,arcsec\,s$^{-1}$ (for the largest, and typically shallowest
maps), this attenuation grows to about 5\% and 15\% at 850 and
450\,\micron, respectively. It is unlikely that this response has had
a significant impact on observations taken to date, since such scans
have been used primarily for shallow cosmology fields for which only
the 850\,\micron\ data are expected to be useful, or Galactic fields
of extended structures (which are less sensitive to this effect). In
the present Starlink \textsc{kapuahi} release there is no option to
de-convolve the system response. However, it is presently being added
and will be present in future releases.

@@

Even after flat-fielding, bolometers may have slightly varying
sensitivities and so the amplitude of the common-mode variations will
also vary from bolometer to bolometer. Comparing each bolometer
time-series with the common-mode signal allows an estimate of the
relative bolometer sensitivity to be obtained. In practice, a least
squares linear fit is performed between the bolometer time-series and
the common-mode to determine a gain [$g_i$ in Eq.~\ref{eq:noise}] and
offset for each bolometer.  The gain and offset for each bolometer is
known as the \model{GAI} model. Each bolometer value can then
optionally be scaled and shifted using these values so that all
bolometers share a common (but as yet unknown) calibration. This
provides an alternative (or additional) flat-fielding strategy to that
described in Section~\ref{sec:bolos}.

@@

Map estimation is accomplished using a nearest-neighbour resampling of
the data onto a pre-defined map grid. For the $i$th map pixel,
$\mathbf{m}(x_i,y_i)$, the brightness is estimated as the weighted
average of the bolometer data samples $\mathbf{b}_j$ that land within
that pixel (from any bolometer or point in time, provided that they
are not flagged as bad or gap-filled, Section~\ref{sec:gaps}),
%
\begin{equation}
  \mathbf{m}(x_i,y_i) = \frac{\sum_j \mathbf{w}_j \mathbf{b}_j }
                             { \sum_j \mathbf{w}_j } .
\end{equation}
%
For the initial iteration the weights $\mathbf{w}_j$ are set to 1, but
subsequently they are set to $1/\sigma_j^2$, the estimated inverse
variance expected from the bolometer white noise levels, as discussed
in Section~\ref{sec:noi}. This weighting scheme is sensible provided
that the bolometer data have no correlated (e.g., low-frequency)
noise.

@@

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{lockman_maps}
\caption{Reduction of a blank-field survey: the Lockman Hole. (a) Raw
  output of SMURF using high-pass filtering as a pre-processing step,
  followed by 4 iterations using only the \model{COM} model to remove
  residual correlated noise. White contours correspond to estimated
  noise levels of 1.25, 2.5 and 5.0 times the minimum noise at the
  centre of the map. (b) The whitened map using a filter based on the
  angular power spectrum of the jackknife map. (c) The whitened map
  cross-correlated with the whitened PSF to identify point sources
  [restricted to a lower-noise region, within the area denoted by the
  second contour of panel (a)]. The image shows the \snr\, with
  3.8-$\sigma$ peaks indicated by blue circles (radius 8\,arcsec). The
  orange ``$+$'' signs show the locations of 1.4\,GHz radio sources
  from \citet{owen2008} with $S\gsim15$\,$\mu$Jy. Of the 10 submm
  peaks, 9 are within a search radius 8\,arcsec of at least one radio
  source. (d) Jackknife map produced from the difference of two maps,
  using the even and odd scan numbers, respectively. Provided that all
  noise sources are statistically uncorrelated between the two halves
  of the data, the map is a plausible realisation of the noise without
  contamination from astronomical sources. (e) The jackknife map
  whitened using the same filter as that in panel (b). The whitened
  jackknife map cross-correlated with the same whitened PSF as in
  panel (c). Again, orange ``$+$'' and blue circles indicate radio
  sources and 3.8-$\sigma$ peaks, respectively. Unlike panel (c),
  there is only a single (apparently) significant peak, and it is not
  in the vicinity of a radio source.}
\label{fig:lockman_maps}
\end{figure*}

@@

In this final example, we analyze a map of M17 which contains bright,
extended emission. The data are from observation 11 on 20110531 using
the 850\,\micron\ array. It is a rotating PONG scan covering a
diameter of 0.375\,deg, with a scan speed of 300\,arcsec\,s$^{-1}$,
and a transverse spacing of 180\,arcsec, taking 37.5\,min to complete.

@@

\includegraphics[width=\linewidth]{cor_pspec_m17_bright_extended}
\end{minipage}
\caption{Lines have same meaning as in Fig.~\ref{fig:m17_def_ps},
except now using the bright extended configuration. This configuration
has resulted in an improved transfer function and \snr\ at large
angular scales. Also note that increasing the filter scale does not
have as large an impact on the small-scale noise as in the baseline
configuration, although it is still noticeably larger up to scales of
about 60\,arcsec.}
\label{fig:m17_be_ps}
\end{figure*}

@@

Maps are produced using the first and second continuous halves of the
data in Fig.~\ref{fig:m17_jk}. This is not an ideal situation, since
the noise properties may evolve with time (e.g., due to changing sky
conditions), leading to a biased estimate of the parent noise
distribution in the complete map from the jackknife. Also, since the
zero-masking depends on the \snr\ of the map, it will be restricted to
regions approximately $\sqrt{2}$ shallower in these maps than for the
full data set. Finally, the cross-linking (positional angles sampled)
is similar, though not identical in the two halves. Ideally one would
have many full maps, as in the case of the Lockman Hole data in the
previous section, from which alternating maps could be combined.

Since our goal in this section is to measure the response of the
map-maker to extended structures, we inject a simulated signal with
power at a range of scales into a relatively empty region of the map.
It is created by drawing a realisation of noise within an 800\,arcsec
on-a-side box, with an angular power spectrum $P(k) \propto k^{-3}$,
which is appropriate for Galactic cirrus clouds
\citep[e.g.,][]{gautier1992}. It is then filtered again with a
14.5\,arcsec Gaussian to model the effect of the \scuba\ optical
response. The RMS of these fluctuations is then normalised to 0.002 pW
so that they are comparable to the dynamic range of M17 itself. The
minimum is then subtracted so that the signal is positive. Finally,
half a period of a (1+cosine)/2 function is used to roll-off the
signal to zero between radii of 300 and 400 arcsec.

@@

The transfer functions $G(k)$ are plotted as thin solid lines in the
right panel of Fig.~\ref{fig:m17_def_ps}. This formula produces
extremely noisy values at large frequencies (small scales), and we
therefore set it to a value of 1 above 0.015\,arcsec$^{-1}$, as well
as any point in the curve where $P_\mathrm{M}(k)$ exceeds
$P_\mathrm{S}(k)$ (i.e., $G(k)$ is assumed to be $\le$1). As expected,
the larger the scale of the filter, the lower the spatial frequency at
which the map transfer function falls. We then correct the map and
jackknife PSDs by dividing by $G(k)$ to produce the thick solid, and
dashed lines in Fig.~\ref{fig:m17_def_ps}b. This shows us that, even
though the raw noise in Fig.~\ref{fig:m17_def_ps}a is lower at small
scales when a smaller-scale filter is used (e.g., the red dashed
line), once we correct for the transfer function, we actually win in a
\snr\ sense at large scales using the larger-scale filter (the
corrected noise is lower), as would be expected.

These tests are then repeated using the bright extended reduction, in
Fig.~\ref{fig:m17_be_ps}. The most obvious improvement with this
reduction over the baseline reduction is that the transfer functions
fall more slowly at large angular scales, accompanied by a slower
increase in noise; in other words, there is greatly improved \snr\ at
large angular scales (an obvious conclusion given the appearance of
the maps in Fig.~\ref{fig:m17_jk}). In fact, using the 900\,arcsec
filter edge, the map response is still above 80 per cent right out to
the largest scale accurately measured in the PSDs, 533\,arcsec, which
is about the largest scale that should be recoverable, given the size
of the array footprint and the fact that we use common-mode rejection.
Another interesting feature of these reductions is that the increase
in small-scale noise as the filter edge is increased is not as drastic
as in the baseline reduction, and the \snr\ improves at scales
$\gsim$60\,arcsec. Finally, note that both the map and jackknife white
noise levels (at scales $\lsim$20\,arcsec) are in excellent agreement,
unlike the previous example.

One case in which the \snr\ is worse using the bright extended
reduction is when using a 150\,arcsec filter. Here the noise is
considerably larger in the bright extended reduction, as evidenced by
the ``kink'' near 150\,arcsec. Referring to the mask contours in the
left panel of the third row in Fig,~\ref{fig:m17_jk}, it is clear
that the map-maker has failed to identify much of the bright, extended
emission in the region of the synthetic source. Each area that is not
within the contours is constrained to zero throughout the solution,
therefore suppressing power (and lowering the transfer function), and
subsequently reducing the \snr\ of the final result. This measurement
serves as a warning: the map-maker response is non-linear when using
\snr\ masking. Harsh filtering can provide misleading results, as in
this example. Maps of faint extended emission will also suffer
considerably, as the structures of interest may not clear the \snr\
threshold for the mask.

@@

The iterative solution is stopped once convergence in the $\chi^2$
statistic, and/or variations in the map itself is achieved. This
enables SMURF to run in a pipeline setting without user interaction
for a wide variety of observations. Furthermore, the execution times
are typically shorter than the observation lengths, and memory
requirements for even the longest \scuba\ observations are within the
capabilities of single, high-end desktop computers. SMURF can
therefore provide real-time feedback at the telescope to observers.

>>>>

There are three general styles of map-making that are relevant to
reducing bolometer data in the literature. The simplest ``direct
methods'' involve some basic processing of the data to remove as much
noise contamination as possible, (e.g., using baseline removal and
other simple filters), and then re-gridding these cleaned data into a
map. Such was the basic recipe for the reduction of chopped data from
SCUBA-2's predecessor SCUBA
\citep{1998ASPC..145..216J,2000ASPC..216..559J}, and MAMBO
\citep[e.g.,][]{omont2001}, another camera from the same generation. A
more recent example is the analysis of SPT data
\citep{schaffer2011}. Generally speaking, such methods are fast,
although depending on the science goals and noise properties of the
data, they may not achieve the best noise performance on the angular
scales of interest. A method for reducing data in fields of faint
point sources with Bolocam \citep[e.g.,][]{laurent2005}, and its
younger sibling the Aztronomical Thermal Emission Camera \citep[AzTEC,
e.g.,][]{scott2008}, is principal component analysis (PCA) cleaning. A
statistical ``black-box'' removes the most correlated components of
the bolometer signals, enabling the detection of point-sources close
to the theoretical white-noise limits of the detectors, with
reasonable computation times when small numbers of bolometers are
involved (hundreds rather than thousands of detectors). However, PCA
cleaning is not a good solution for producing maps of extended
structures, since such sources produce correlated signals amongst many
detectors, and are removed by this procedure \citep[an exception is
the iterative PCA approach of][]{aguirre2011}. Furthermore, in the
case of \scuba, performing PCA on even a single subarray (typically
$\sim$900 detectors) can be prohibitively slow.

@@

The third approach adopted here for \scuba\ is a compromise between
the previous two methods. Under the assumption that a significant
portion of the (predominantly low-frequency) non-white noise sources
can be modelled, an iterative solution is obtained for both the
astronomical image and the parameters of the noise model. Since the
remaining (non-modelled) noise sources are assumed to be white, a
single scalar \rms\ may be calculated for all of the data points from
a given bolometer to characterise its noise distribution (since the
noise at any instant in time is uncorrelated with others, and with the
data from other bolometers), greatly simplifying the inversion step
that is so complicated in the maximum-likelihood methods. In the
iterative approach memory usage scales linearly with $N$, the product
of the number of bolometers, $N_\mathrm{b}$, with the number of
samples in time, $N_\mathrm{t}$. The computation time is the product
of the number of iterations, $N_\mathrm{iter}$, with the time per
iteration, $t_\mathrm{iter}$. For the algorithm described in this
paper, $N_\mathrm{iter}$ is typically in the range $\sim$5--100
(depending on the \snr\ of large-scale structures, and the size of the
map), and $t_\mathrm{iter}$ scales as $N_\mathrm{t} \log
N_\mathrm{t}$, since Fast Fourier Transforms (FFTs) of the time-series
are typically performed. By comparison, the fastest maximum-likelihood
methods that account for inter-bolometer correlations have an
execution time with an $N_\mathrm{b}^2$ dependence, although it is
possible to limit memory usage to be linear in $N_\mathrm{t}$
\citep[e.g.,][]{patanchon2008}.

@@

A reasonable model for correlated noise in \scuba\ is a single
``common-mode'' signal seen by all of the bolometers. Iterative
estimation and removal of this signal significantly lowers the $1/f$
knee, without compromising structures on angular scales smaller than
the array footprint. Residual independent drifts at lower frequencies
are removed with an iterative high-pass filter. This strategy enables
SMURF to reduce data on time scales commensurate with the observing
times (or significantly faster), on single high-end desktop
computers. For reference, all of the data analysis in this paper was
performed on a machine with a 64-bit 8-core central processing unit
operating at 2.7\,GHz, and 48\,Gb of memory. SMURF has been
successfully used as part of a real-time pipeline \citep[based on the
Observatory Reduction and Acquisition Control - Data Reduction system,
or ORAC-DR,][]{gibb2005,2008AN....329..295C} offering feedback to
observers at the telescope. The pipeline is also used to generate
products for the JCMT Science Archive \citep{2011ASPC..442..203E}
hosted by the Canadian Astronomy Data Centre (CADC).

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{skynoise}
\caption{Illustration of the dominance of atmospheric variations on
  long-timescales using 35\,min. of data taken in highly-variable
  weather conditions (line-of-sight 450\,\micron\ opacity ranging from
  3.6 to 7.7) from the fully-commissioned instrument, observation 18
  on 20110324. Red and blue lines show the 850 and 450\,\micron\
  common-mode signals, corresponding to the left- and right- vertical
  axes, respectively. The black line shows the line-of-sight opacity
  from the WVM as a proxy for atmospheric emission (arbitrary
  scale). Bolometer baseline drifts on timescales $\gsim$100\,s are
  clearly correlated with the atmosphere in this case.}
\label{fig:skynoise}
\end{figure}

The data shown in Figs.~\ref{fig:bolos_mix} and \ref{fig:pspec} were
taken during fairly stable weather conditions (although with different
450\,\micron\ line-of-sight opacities $\sim$0.9 and 6.5 in the S2SRO
and post-upgrade data sets, respectively), and are dominated at
low-frequencies ($\lsim$2\,Hz) by scan-synchronous noise, rather than
underlying variations in the atmosphere, or ``sky noise''. However,
only $\sim$100\,s of data were analysed. In Fig.~\ref{fig:skynoise},
an order-of-magnitude longer (fully-commissioned) data set is plotted
from observation 18 on 20110324, during highly variable weather
conditions (450\,\micron\ line-of-sight opacities ranging from 3.6 to
7.7, scaled from the WVM data which are shown as a black line). Unlike
the earlier plots, it is clear that the bolometer baseline drifts on
timescales $\gsim$100\,s are tightly correlated with the atmospheric
opacity, which serves as a proxy for atmospheric emission. The weaker
correlation between the WVM and the 450\,micron\ bolometers is
probably due to the atmosphere being more transmissive at 850\micron,
therefore the emission detected at 450\,\micron\ comes from water
vapour that is closer to the telescope.

@@

As an alternative, we also investigated an algorithm in which the FFT
of $x_i$ is simply truncated to the target sample rate before
transforming back to the time domain (i.e., applying a hard-edged
low-pass filter). In practice the noise performance was
indistinguishable, and required slightly longer execution time
compared with the previous method, and is therefore not used.

@@

SMURF uses FFTs of the bolometer data extensively for filtering. Data
that have been flagged as bad for any reason (for instance, due to the
presence of spikes, steps, or unusual common-mode signal) need to be
excluded. For this reason, each contiguous block of bad data samples
is filled with artificial data before taking the FFT. A least squares
linear fit is performed to the 50 samples preceding the block, and a
similar fit is performed to the 50 samples following the block. These
are used to estimate the expected values at the start and end of the
block of bad values. The bad values are then replaced by linear
interpolation between the expected start and end values. Gaussian
noise is added with a standard deviation equal to the mean of the RMS
residuals in the two fits. These flagged and filled portions of the
data are then given a weight of zero when estimating the map.

@@

Finally, data that are taken while the telescope is either moving
extremely slowly, or extremely fast, are flagged and ignored when
estimating the map. When the telescope is moving slowly, astronomical
signals appear at low frequencies in the bolometer data that are
dominated by $1/f$ noise (in the extreme case of no motion, the
bolometers have zero sensitivity). When the telescope is moving
significantly faster than 600\,arcsec\,s$^{-1}$, both the thermal
response of the bolometers, and the anti-aliasing filter in the
readout electronics (Section~\ref{sec:bolos}) can significantly
distort signals.  The default allowable range of speed is
30--1000\,arcsec\,s$^{-1}$. These thresholds primarily reject short
periods when the telescope was stationary prior to commencing the scan
pattern (generally speaking the speed is constant once the scan
begins). The scan speed can also briefly exceed the upper limit when
approaching the maximum requested speed of 600\,arcsec as the pointing
system struggles at high elevation.

@@

In principle, the effects of the bolometer thermal response,
anti-aliasing filter, and data acquisition lags should be accounted
for; together they produce a net lag in the bolometer signals of about
4.5\,ms with respect to the pointing data
(Section~\ref{sec:bolos}). This can be modelled by convolving the
``pure'' bolometer signals with a single system response
function. Similarly, this effect could be removed using de-convolution
as a pre-processing step. The impact of this lag is primarily a
reduction in point-source sensitivity since scans at different
position angles will detect the source at slightly different positions
(and therefore reduce the peak signal once averaged together). This
attenuation is most significant at 450\,\micron\ given the smaller
beam size, and at the greatest scan speeds. For reference, at a speed
of 150\,arcsec\,s$^{-1}$ (typical for CV daisies of calibrators and
deep fields) the attenuation is a negligible $<$1\% at 850\,\micron,
and $\sim$1\% at 450\,\micron. However, at the maximum speed of
600\,arcsec\,s$^{-1}$ (for the largest, and typically shallowest
maps), this attenuation grows to about 5\% and 15\% at 850 and
450\,\micron, respectively. It is unlikely that this response has had
a significant impact on observations taken to date, since such scans
have been used primarily for shallow cosmology fields for which only
the 850\,\micron\ data are expected to be useful, or Galactic fields
of extended structures (which are less sensitive to this effect). In
the present Starlink \textsc{kapuahi} release there is no option to
de-convolve the system response. However, it is presently being added
and will be present in future releases.

@@

Even after flat-fielding, bolometers may have slightly varying
sensitivities and so the amplitude of the common-mode variations will
also vary from bolometer to bolometer. Comparing each bolometer
time-series with the common-mode signal allows an estimate of the
relative bolometer sensitivity to be obtained. In practice, a least
squares linear fit is performed between the bolometer time-series and
the common-mode to determine a gain ($g_i$ in Eq.~\ref{eq:noise}) and
offset for each bolometer.  The gain and offset for each bolometer is
known as the \model{GAI} model. Each bolometer value can then
optionally be scaled and shifted using these values so that all
bolometers share a common (but as yet unknown) calibration. This
provides an alternative (or additional) flat-fielding strategy to that
described in Section~\ref{sec:bolos}.

@@

Map estimation is accomplished using a nearest-neighbour resampling of
the data onto a pre-defined map grid. For the $i$th map pixel,
$\mathbf{m}(x_i,y_i)$, the brightness is estimated as the weighted
average of the bolometer data samples $\mathbf{b}_j$ that land within
that pixel (from any bolometer or point in time, provided that they
are not flagged as bad or gap-filled, Section~\ref{sec:gaps}),
%
\begin{equation}
  \mathbf{m}(x_i,y_i) = \frac{\sum_j \mathbf{w}_j \mathbf{b}_j }
                             { \sum_j \mathbf{w}_j } .
\end{equation}
%
For the initial iteration the weights $\mathbf{w}_j$ are set to 1, but
subsequently they are set to $1/\sigma_j^{\mathbf{w}2}$, the estimated
inverse variance expected from the bolometer white noise levels, as
discussed in Section~\ref{sec:noi}. This weighting scheme is sensible
provided that the bolometer data have no correlated (e.g.,
low-frequency) noise.

@@

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{lockman_maps}
\caption{Reduction of a blank-field survey: the Lockman Hole. (a) Raw
  output of SMURF using high-pass filtering as a pre-processing step,
  followed by 4 iterations using only the \model{COM} model to remove
  residual correlated noise. White contours correspond to estimated
  noise levels of 1.25, 2.5 and 5.0 times the minimum noise at the
  centre of the map. (b) The whitened map using a filter based on the
  angular power spectrum of the jackknife map. (c) The whitened map
  cross-correlated with the whitened PSF to identify point sources
  [restricted to a lower-noise region, within the area denoted by the
  second contour of panel (a)]. The image shows the \snr\, with
  3.8-$\sigma$ peaks indicated by blue circles (radius 8\,arcsec). The
  orange ``$+$'' signs show the locations of 1.4\,GHz radio sources
  from \citet{owen2008} with $S\gsim15$\,$\mu$Jy. Of the 10 submm
  peaks, 9 are within a search radius of 8\,arcsec of at least one
  radio source. (d) Jackknife map produced from the difference of two
  maps, using the even and odd scan numbers, respectively. Provided
  that all noise sources are statistically uncorrelated between the
  two halves of the data, the map is a plausible realisation of the
  noise without contamination from astronomical sources. (e) The
  jackknife map whitened using the same filter as that in panel
  (b). The whitened jackknife map cross-correlated with the same
  whitened PSF as in panel (c). Again, orange ``$+$'' and blue circles
  indicate radio sources and 3.8-$\sigma$ peaks, respectively. Unlike
  panel (c), there is only a single (apparently) significant peak, and
  it is not in the vicinity of a radio source.}
\label{fig:lockman_maps}
\end{figure*}

@@

In this final example, we analyse a map of M17 which contains bright,
extended emission. The data are from observation 11 on 20110531 using
the 850\,\micron\ array. It is a rotating PONG scan covering a
diameter of 0.375\,deg, with a scan speed of 300\,arcsec\,s$^{-1}$,
and a transverse spacing of 180\,arcsec, taking 37.5\,min to complete.

@@

\includegraphics[width=\linewidth]{cor_pspec_m17_bright_extended}
\end{minipage}
\caption{Lines have same meaning as in Fig.~\ref{fig:m17_def_ps},
except now using the bright extended configuration. This configuration
has resulted in an improved transfer function and \snr\ at large
angular scales. Also note that increasing the filter scale does not
have as large an impact on the small-scale noise as in the baseline
configuration.}
\label{fig:m17_be_ps}
\end{figure*}

@@

Maps are produced using the first and second continuous halves of the
data in Fig.~\ref{fig:m17_jk}. This is not an ideal situation, since
the noise properties may evolve with time (e.g., due to changing sky
conditions), leading to a biased estimate of the parent noise
distribution in the complete map from the jackknife. Also, since the
zero-masking depends on the \snr\ of the map, it will be restricted to
regions approximately $\sqrt{2}$ shallower in these maps than for the
full data set. Finally, the cross-linking (position angles sampled) is
similar, though not identical in the two halves. Ideally one would
have many full maps, as in the case of the Lockman Hole data in the
previous section, from which alternating maps could be combined.

Since our goal in this section is to measure the response of the
map-maker to extended structures, we inject a simulated signal with
power at a range of scales into a relatively empty region of the map.
It is created by drawing a realisation of noise from an angular power
spectrum $P(k) \propto k^{-3}$, which is appropriate for Galactic
cirrus clouds \citep[e.g.,][]{gautier1992}, within an 800\,arcsec
on-a-side box. It is then filtered again with a 14.5\,arcsec Gaussian
to model the effect of the \scuba\ optical response. The RMS of these
fluctuations is then normalised to 0.002 pW so that they are
comparable to the dynamic range of M17 itself. The minimum is then
subtracted so that the signal is positive. Finally, multiplication by
half a period of a (1+cosine)/2 function is used to roll-off the
signal to zero between radii of 300 and 400 arcsec.

@@

The transfer functions $G(k)$ are plotted as thin solid lines in
Fig.~\ref{fig:m17_def_ps}b. This formula produces extremely noisy
values at large frequencies (small scales), and we therefore set it to
a value of 1 above 0.015\,arcsec$^{-1}$, as well as any point in the
curve where $P_\mathrm{M}(k)$ exceeds $P_\mathrm{S}(k)$ [i.e., $G(k)$
is assumed to be $\le$1]. As expected, the larger the scale of the
filter, the lower the spatial frequency at which the map transfer
function falls. We then correct the map and jackknife PSDs by dividing
by $G(k)$ to produce the thick solid, and dashed lines in
Fig.~\ref{fig:m17_def_ps}b. This shows us that, even though the raw
noise in Fig.~\ref{fig:m17_def_ps}a is lower at small scales when a
smaller-scale filter is used (e.g., the red dashed line), once we
correct for the transfer function, we actually win in a \snr\ sense at
large scales using the larger-scale filter (the corrected noise is
lower), as would be expected.

These tests are then repeated using the bright extended reduction, in
Fig.~\ref{fig:m17_be_ps}. The most obvious improvement with this
reduction over the baseline reduction is that the transfer functions
fall more slowly at large angular scales, accompanied by a slower
increase in noise; in other words, there is greatly improved \snr\ at
large angular scales (an obvious conclusion given the appearance of
the maps in Fig.~\ref{fig:m17_jk}). In fact, using the 900\,arcsec
filter edge, the map response is still above 80 per cent right out to
the largest scale accurately measured in the PSDs, 533\,arcsec, which
is about the largest scale that should be recoverable, given the size
of the array footprint and the fact that we use common-mode rejection.
Another interesting feature of these reductions is that the increase
in small-scale noise as the filter edge is increased is not as drastic
as in the baseline reduction. Finally, note that both the map and
jackknife white noise levels (at scales $\lsim$20\,arcsec) are in
excellent agreement, unlike the previous example.

One case in which the \snr\ is worse using the bright extended
reduction is when using a 150\,arcsec filter. Here the noise is
considerably larger in the bright extended reduction, as evidenced by
the ``kink'' near 150\,arcsec. Referring to the mask contours in the
left panel of the third row in Fig,~\ref{fig:m17_jk}, it is clear that
the map-maker has failed to identify much of the bright, extended
emission in the region of the synthetic source. Each area that is not
within the contours is constrained to zero throughout the solution,
therefore suppressing power (and lowering the transfer function), and
subsequently reducing the \snr\ of the final result. This measurement
serves as a warning: the map-maker response is non-linear when using
\snr\ masking. Harsh filtering can provide misleading results, as in
this example. Maps of faint extended emission will also suffer
considerably, as the structures of interest may lie below the \snr\
threshold for the mask.

@@

The iterative solution is stopped once convergence in the map itself
is achieved. This enables SMURF to run in a pipeline setting without
user interaction for a wide variety of observations. Furthermore, the
execution times are typically shorter than the observation lengths,
and memory requirements for even the longest \scuba\ observations are
within the capabilities of single, high-end desktop computers. SMURF
can therefore provide real-time feedback at the telescope to
observers.
